---
title: "dplyr"
output: html_document
---

Andrew Innes

9/9/17

## Filtering Columns and Rows

When using dplyr, you can filter columns and rows by using select and filter.  Let's look at an example using the Lahman baseball database.  We first have to load the Lahman and dplyr packages.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(dplyr)
```

Now, suppose we would like to see the homerun totals for the 1927 New York Yankees.  We could run the following code:

```{r warning=FALSE}
Batting%>%
  select(playerID,yearID,teamID,HR)%>%
  filter(teamID=='NYA' & yearID==1927)
```

#----------------------------------------

Now we would like to see players from the New York Yankees that hit 40 or more homeruns in a given year. We could run the following code:

```{r}
Batting %>%
  select (playerID, yearID, teamID, HR) %>%
  filter (HR>=40 & teamID=='NYA')
```

#-----------------------------------------

We can take the same code and switch the filter and select lines.  It does not matter the order it is in.

```{r}
Batting%>%
  filter (HR>=40 & teamID=='NYA')%>%
  select (playerID, yearID, teamID, HR)
```

#-----------------------------------------

Next we would like to see players that hit over 40 homeruns but had less than 50 strikeouts in a single season.  We could run the following code:

```{r}
Batting%>%
  select(playerID,yearID,teamID,HR,SO)%>%
  filter(HR>40 & SO<60)
```

#-----------------------------------------

Now we want to see the players from the Phillies that hit over 30 homeruns in the 1970s. We could run the following code:

```{r}
Batting%>%
  select(playerID,yearID,teamID,HR)%>%
  filter(HR>30 & teamID=='PHI' & yearID>=1970 & yearID<=1979)
```

#---------------------------------------------

#Using Arrange Function
We would like to see players who hit more than 50 homeruns in a single season with the highest hitting homerun scorers at the top of our list.  To do this we need the arrange function in descending order.  So we could enter the code as follows:

```{r}
Batting%>%
  select(playerID,yearID,teamID,HR)%>%
  filter(HR>50)%>%
  arrange(desc(HR))
```

#-------------------------------------

Next we want to arrange players by least amount of strikeouts.  We want to see what players hadd less than 10 strikeouts but more than 400 at bats.  We could use the following code:

```{r}
Batting%>%
  select(playerID,yearID,teamID,SO,AB)%>%
  filter(SO<10 & AB>=400)%>%
  arrange(SO)
```

#------------------------------------

#Using Grouping
In order to see how many homeruns a player has hit during their whole career we use the group_by function.  In this instance we look at the homerun totals for Babe Ruth's career.

```{r}
Batting%>%
  filter(playerID=='ruthba01')%>%
  group_by(playerID)%>%
  summarize(career_HR=sum(HR))
```

#---------------------------------

Next we decide to look at career homerun totals for players who hit over 600 homeruns in order from most to least.

```{r}
Batting%>%
  group_by(playerID)%>%
  summarize(career_HR=sum(HR))%>%
  filter(career_HR>=600)%>%
  arrange(desc(career_HR))
```

#---------------------------------

We want to see the average homerun totals greater than or equal to 30 per season.  We can write the code as follows:

```{r}
Batting%>%
  group_by(playerID)%>%
  summarize(average_HR=mean(HR))%>%
  filter(average_HR>=30)%>%
  arrange(desc(average_HR))
```

#---------------------------------

Finally we want to see any player who has hit over 50 homeruns in a single season after the 1970s.  But we only want a players name to display once.  We could write code as follow:

```{r}
Batting%>%
  filter(yearID>=1970)%>%
  group_by(playerID)%>%
  summarize(maxHR=max(HR))%>%
  filter(maxHR>50)%>%
  select(playerID)
```

